Corotating Dust Clumps Around Adolescent M-Dwarfs: Four Years of TESS

### 2023-06-26 paper/ms.tex
update ms

\newcommand{\rprs}{{$R_p/R_{\star}$}}
\newcommand{\vsini}{{$V \sin i$}}
\newcommand{\teff}{T$_{\textrm{eff}}$}
\newcommand{\kms}{{km\,s$^{-1}$}}
\newcommand{\gcc}{{g\,cm$^{-3}$}}
\newcommand{\rstar}{{$R_\star$}}
\newcommand{\rhostar}{{$\rho_\star$}}
\newcommand{\mearth}{{M$_\oplus$}}
\newcommand{\rearth}{{R$_\oplus$}}
\newcommand{\rsun}{{R$_\odot$}}
\newcommand{\msun}{{M$_\odot$}}

@@

\correspondingauthor{TBD}
\email{TBD}

@@

\shortauthors{TBD}

@@

\author{Authors to include (ORDER TBD)}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are pre-main-sequence
  M-dwarfs with almost-periodic optical modulation that includes both
  sharp and broad troughs.  The troughs are usually superposed over
  smooth, spot-like modulation.  These objects are typically 10-100
  million years (Myr) old;  they do not have primordial gas-rich
  disks.  Nonetheless, the preferred explanations for their optical
  variability invoke long-lived material in orbit at the Keplerian
  corotation radius, most likely transiting dust clumps.
  Here, we report new CQVs discovered in TESS data
  collected between July~2018 and Sep~2022.  From our analysis of
  \nstarssearched\ M-dwarfs with $T$$<$16 and $d$$<$150\,pc, we find
  \ngoods\ gold-standard CQVs.  Most of these discoveries are new, and
  they include the brightest and closest examples of this class of
  object known ($T$$\approx$11; $d$$\approx$20\,pc).  A few objects
  are outliers among outliers; LP 12-502 for instance shows a ``dip
  complex'' with a total duration that is nearly constant over more
  than 1000 cycles, but which exhibits anywhere from four to eight
  local minima per cycle.  LP-502 also demonstrated drastic shape
  changes over less than one cycle, and (locally) shows up to four distinct
  periods.  More generally, we find that none of the CQVs maintain a
  fixed light curve shape over timescales of more than a few hundred
  cycles -- they are all {\it quasi}periodic.  Our sample will
  help facilitate future modeling and observational efforts aimed at understanding
  these objects, and connecting them to the broader
  contexts of star, disk, and
  exoplanet evolution.
\end{abstract}

@@

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets.

Data from K2 and TESS have yielded a new class of variable star whose
root cause is not yet qualitatively understood: complex quasiperiodic
variables (CQVs).  Figure~\ref{fig:f1} shows a few examples.  These
objects are identified phenomenologically using their optical light
curves, which show quasiperiodic troughs that can be either sharp or
broad, often superposed on smooth spot-like modulation.  Some of the
objects show up to eight local minima (dips) per cycle; sharp peaks
are also common.  Most of these variable stars turn out to be
pre-main-sequence M-dwarfs without near-infrared excesses, with ages
of $\approx$5-150 million years (Myr), and rotation periods of at most
two days; they are observed to be $\approx$1\% of M-dwarfs younger
than 100 Myr
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2022AJ....163..144G}.
The dips can be chromatic, with a reddening law plausibly consistent
with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

%TODO rewrite, but favor dust clumps even more
The issue of what causes the variability that defines the CQVs has yet
to be resolved.  There are a few competing explanations.  All young
M-dwarfs are spotted, which produces flux variations over
characteristic timescales of the rotation period, $P_{\rm rot}$, and
its half-harmonic, $0.5\,P_{\rm rot}$.  The observed dips occur over
durations as short as $0.05\,P_{\rm rot}$; a ``starspot-only''
scenario is therefore highly unlikely for any of the objects with
sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The more likely
scenarios instead invoke sharp geometries with material extrinsic to
the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  In the
``clump'' scenario, opaque dust clumps orbiting near the Keplerian
corotation radius eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  In the
``prominence'' scenario, long-lived condensations of plasma are
trapped along the star’s magnetic field lines, also near corotation
\citep{2022MNRAS.514.5465W}.  And in a ``screen'' scenario, the inner
wall of a quiescent circumstellar disk blocks a portion of the stellar
surface to produce sudden dips whenever spots come into view
\citep{2019ApJ...876..127Z}.  In our view, the arguments seem best
established for the dust clump and prominence hypotheses
\citep{2023MNRAS.518.4734S,2022MNRAS.514.5465W}.  However, unambiguous
evidence has yet to be presented to demonstrate which, if any, of
these scenarios is correct.

It has not yet been possible to rule between the candidate mechanisms
CQVs have historically been both hard to discover and hard to
characterize.   Discoverability has been linked to rarity: CQVs
comprise about 1\% of the youngest 1\% of M-dwarfs
\citep{2018AJ....155..196R}.  Out of the millions of stars monitored
by K2 and TESS, about 50 CQVs have been reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2022AJ....163..144G,2023ApJ...945..114P}.
The known CQVs are correspondingly faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  This renders time-series optical
spectroscopy at high resolution technically impossible, despite its
potential utility in ruling between the models.

@@

A point on nomenclature.  CQVs have been called ``transient and
persistent flux dips'', ``scallop shells'', ``batwings'',
\citep{2017AJ....153..152S,2018AJ....155...63S} ``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  The
CQVs are not ``dippers'', which are classical T Tauri stars with
infrared excesses whose optical variability is linked to obscuring
inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  At the risk of
introducing yet another
standard,\footnote{\url{https://xkcd.com/927/}} we prefer to make the
technical distinction that the CQVs are, over timescales of more than
a few cycles, not exactly periodic.  Their shapes evolve, usually over
some decoherence timescale, similar to variability induced by surface
inhomogeneities such as starspots.  While the three-type
classification scheme proposed by \citet{2017AJ....153..152S} may
indeed provide some helpful visual distinctions among the CQV class,
it seems likely (we hope!) that they are all explained by a single
underlying phenomenon, and so we opt to refer to them by a single
empirically descriptive name.

@@

		\vspace{-0.4cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{models.pdf}
		}
	\end{center}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
		{\it Top:} Phase-folded TESS light curves of three CQVs.  Each is
		stacked over one month.  Gray are raw 2-minute data; black bins to
		300 points per cycle.  Periods in hours are in the bottom right of
		each panel.  In order left-to-right, the objects are LP 12-502
		(TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and TIC
		425933644 (Sector~28).
		{\it Bottom:} Plausible cartoon models for the phenomenon.  The
		``clump'' or ``prominence'' scenarios seem most plausible, given
		the stability of the phenomenon and the lack of observed infrared
		excesses.
	}
	\label{fig:f1}
\end{figure*}

@@

To assess the completeness of the resulting 120-second cadence data
that we analyze, we cross-matched TIC8 \citep{2018AJ....156..102S}
against the Gaia DR2 point-source catalog \citep{2018A&A...616A...1G}.
We opted for Gaia DR2 rather than DR3 because the base catalog for
TIC8 was Gaia DR2, which facilitated a one-to-one crossmatch using the
Gaia source identifiers.  This exercise showed that for $T$$<$16
M-dwarfs, the TESS 2-minute data are roughly 50\% complete at
$\approx$50\,pc.  At $<$20\,pc, $\gtrsim$80\% of the $T$$<$16 M-dwarfs
have at least one sector of short-cadence data; at $>$100\,pc,
$\lesssim$10\% of such M-dwarfs have at least one sector of
short-cadence data.  Armed with this understanding, we then used our
cross-match between Gaia DR2 and TIC8 to select our stars of interest,
which we defined as stars with 120-second cadence TESS light curves
that satisfied
\begin{align}
  T &< 16 \quad&(\mathrm{Bright\ for\ TESS})\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only}) \\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})  \\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}),
\end{align}
for $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ the Gaia $G$-band
absolute magnitude, $\varpi_{\rm as}$ the parallax in units of
arcseconds, and a geometric distance $d$ defined by inverting the
parallax and ignoring any zero-point correction.  This selection
function includes dwarf stars later than spectral types of
$\approx$K6V.  This target sample includes \nstarssearched\ stars with
approximately M-dwarf spectral type, down to $T$$<$16 and out to
$d$$<$150\,pc.

@@

Previous methods that have yielded CQV discoveries include visually
examining stars known to be in young clusters
\citep{2016AJ....152..114R,2017AJ....153..152S}, and automatically
flagging rapid rotators with a large number of strong Fourier
harmonics \citep{2019ApJ...876..127Z}.  The latter approach requires
visual vetting, since most rapid rotators with many Fourier harmonics
are false positives such as eclipsing binaries or multiple stars
blended into a single photometric aperture.  In this work, we
implemented a new search approach based on counting the number of
sharp local minima in phase-folded light curves, while also using the
previously tested Fourier approach.  We applied these two search
techniques independently.

{\bf todo for Rahul or Saul: explain the approach, in a few
paragraphs.  Was the SAP\_FLUX or PDCSAP used? etc. }

@@

@@

\paragraph{Ages}
We estimated the stellar ages by making probabilistic spatial and
kinematic associations between the CQVs and known clusters in the
solar neighborhood.  For most of the stars in our sample, we were able
to do this using BANYAN\,$\Sigma$
\citep{2018ApJ...856...23G}.\footnote{\url{https://github.com/jgagneastro/banyan_sigma},
git commit \texttt{394b486}} This algorithm calculates the probability
that a given star belongs to one of 27 young clusters (or
``assocations'') within 150\,pc of the Sun, by modeling the clusters
as multivariate Gaussians in 3-D position and 3-D velocity space.  We
used the Gaia DR2 sky positions, proper motions, and distances to
calculate the membership probabilities.  BANYAN\,$\Sigma$ in turn
analytically marginalizes over the radial velocity dimension.  The
probabilities returned by this procedure are qualitatively useful, but
quantitatively dubious due to the non-Gaussian nature of most groups
within the solar neighborhood \citep[see
e.g.][Figure~10]{2021ApJ...917...23K}.

@@

We determine the stellar effective temperature and radii through SED
fitting; the masses are then determined based on interpolation against
the ``magnetic'' models of \citet{2016A&A...593A..99F}.

For the SED fitting, we use \texttt{astroARIADNE}
\citep{2022MNRAS.513.2719V}, tuned to use only the BT-Settl stellar
atmosphere models \citep{Allard2012} based on the
\citet{2009ARA&A..47..481A} solar abundances.  The broadband
magnitudes we considered included $GG_{\rm BP}G_{\rm RP}$ from Gaia
DR2, $Vgri$ from APASS, $JHK_{\rm S}$ from 2MASS, $riz$ from
Pan-STARRS1, SDSS $riz$, and the WISE 1-2 passbands.  We specifically
omitted UV flux measurements to avoid biasing our fit with any
possible chromospheric UV excess.
\texttt{astroARIADNE} compares the measured broadband flux measurements
against pre-computed model grids, and by default fits for six
parameters:
$\{ T_{\rm eff}, R_\star, A_{\rm V}, \log g, [{\rm Fe/H}], d \}$.
The distance  prior is drawn from \citet{2021AJ....161..147B}.
The surface gravity and metallicity are generally unconstrained.
And finally, given our particular use-case, 
we assumed the following priors for the temperature, stellar size, and
extinction:
\begin{align}
  T_{\rm eff} / {\rm K}    &\sim \mathcal{N}(3000, 1000), \\
  R_\star / {\rm R}_\odot  &\sim \mathcal{T}_{\rm N}(0.5, 0.3, 0.1, 1.5), \\
  A_{\rm V} / {\rm mag}    &\sim \mathcal{U}(0, 0.2),
\end{align}
for $\mathcal{N}$ the Gaussian and $\mathcal{U}$ the uniform
distributions, and $\mathcal{T}(\mu, \sigma, a, b)$ a truncated normal
distribution with mean $\mu$, standard deviation $\sigma$, and lower
and upper bounds $a$ and $b$.
Using \texttt{Dynesty} \citep{2020MNRAS.493.3132S},
we statically sampled the posterior probability assuming the default
Gaussian likelihood, and set a stopping threshold of ${\rm d}\log
\mathcal{Z} < 0.01$, for $\mathcal{Z}$ the evidence.

Given the effective temperatures and stellar radii from the SED fit,
we then estimate the mass by interpolating against the ``magnetic''
models of \citet{2016A&A...593A..99F}.  The need for models that
incorporate some form of magnetic correction is well-documented for
our regime of active, young M-dwarfs
\citep[e.g.][]{2012ApJ...756...47S,2015ApJ...804..146D,2020ApJ...891...29S}.
While one could of course allow a range of magnetic field strengths,
and a corresponding range of fractions for starspot converage
\citep{2020ApJ...891...29S}, we prefer the magnetic
\citet{2016A&A...593A..99F} models based on the simplicity of their
assumption that the surface magnetic field strength will be fixed at
thermal equipartition.
We therefore interpolate our SED-measured temperatures and radii, at
the known cluster-based stellar ages ages, in order to estimate
(model-dependent) stellar masses.
For the two systems without known ages (see below), we omit this mass
estimate, due to the degeneracies in the inference problem.

@@

\begin{figure*}[!t]
	\begin{center}
    \centering
    \includegraphics[width=0.96\textwidth]{402980664_S0018_120sec_cpvvetter.pdf}
		\vspace{-0.45cm}
		\caption{
      {\bf Validation plots used to label CQVs}.  The complete figure
      set of \ncpvsfound\ images is available online.
      Panels are as follows.
      {\it a)}: Phase-folded light curve; gray points are raw 2-minute
      data and black points are binned to 200 points per cycle.
      {\it b)}: Phase-dispersion minimization (PDM) periodogram.
      Dotted lines show up to the 10$^{\rm th}$ harmonic and
      subharmonic.
      {\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
      circles displayed for scale.
      {\it d)}: Cleaned light curve, binned to 20-minute cadence, in
      Barycentric TESS Julian Date (BTJD).
      {\it e)}: Phase-folded light curve, binned to 100 points per
      cycle.  The gray line denotes the automated spline-fit to the
      wrapped phase-folded light curve, and small gray triangles
      denote automatically identified local minima.
      {\it f)}: Phase-folded light curve at twice the peak period.
      {\it g)}: Phase-folded light curve at half the peak period.
      {\it h)}: Phase-folded time-series within the ``background''
      aperture defined in the SPOC light curves.
      {\it i)}: Phase-folded flux-weighted centroid in the column
      direction.
      {\it j)}: Phase-folded flux-weighted centroid in the row
      direction.
      {\it k)}: Gaia DR2 color--absolute magnitude diagram.     
      {\it l)}: Information from Gaia DR2, TIC8, and the automated
      dip-counting search pipeline.  ``Neighbors'', abbreviated
      ``nbhr'', are listed within apparent distances of 2 TESS pixels
      if $\Delta T$$<$2.5.
      {\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
      using positions, proper motions, and the parallax.
      }
		\label{fig:vet}
	\end{center}
\end{figure*}

\section{LP 12-502}

>>>>

\newcommand{\rprs}{{$R_p/R_{\star}$}}
\newcommand{\vsini}{{$V \sin i$}}
\newcommand{\kms}{{km\,s$^{-1}$}}
\newcommand{\gcc}{{g\,cm$^{-3}$}}
\newcommand{\rstar}{{$R_\star$}}
\newcommand{\rhostar}{{$\rho_\star$}}
\newcommand{\mearth}{{M$_\oplus$}}
\newcommand{\rearth}{{R$_\oplus$}}
\newcommand{\rsun}{{R$_\odot$}}
\newcommand{\msun}{{M$_\odot$}}

\newcommand{\teff}{T_{\rm eff}}
\newcommand{\ms}{M_\star}
\newcommand{\rs}{R_\star}

@@

%\title{Complex Quasiperiodic Variables From Four Years of TESS}

\correspondingauthor{Luke G. Bouma}
\email{luke@astro.caltech.edu}

@@

\shortauthors{Bouma, Jayararaman, et al.}

@@

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation that
  is likely induced by long-lived transiting dust clumps in orbit at
  the Keplerian corotation radius.  Here, we report new CQVs
  discovered in TESS data collected between July~2018 and Sep~2022.
  From our analysis of \nstarssearched\ K- and M-dwarfs with $T$$<$16
  and $d$$<$150\,pc, we find \ngoods\ gold-standard CQVs.  Most of
  these discoveries are new, and they include the brightest and
  closest examples of this class of object known ($T$$\approx$11;
  $d$$\approx$20\,pc).  A few objects are outliers among outliers; LP
  12-502 for instance shows a ``dip complex'' with a period and total
  duration that are fixed over more than 1000 cycles, but in detail
  this object shows anywhere from four to eight local minima per
  cycle.  LP-502 also demonstrated drastic shape changes over less
  than one cycle, and at times displayed distinct superposed periods
  simultaneously.  Generally speaking, we find that none of the CQVs
  maintain a fixed light curve shape over timescales of more than a
  few hundred cycles -- they are all {\it quasi}periodic.  We revisit
  the arguments for why corotating dust clumps are the most likely
  explanation, though there are other possibilities, and all lines of
  evidence are currently indirect.  In the future, we expect that our
  sample will help facilitate modeling and observational efforts aimed
  at understanding these objects, and connecting them to the broader
  contexts of star, disk, and exoplanet evolution.
\end{abstract}

@@

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep{2010exop.book...55W}.

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show quasiperiodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle; sharp peaks
are also common.  Most CQVs are pre-main-sequence M-dwarfs without
near-infrared excesses, with ages of $\approx$5-150 million years
(Myr), and rotation periods of at most two days; they are observed to
be $\approx$1-3\% of M-dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  The dips can be
chromatic, with a reddening law plausibly consistent with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with a few
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The more likely
scenarios invoke sharp geometries with material extrinsic to the
stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  In the
``clump'' scenario, opaque dust clumps orbiting near the Keplerian
corotation radius eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  In the
``prominence'' scenario, long-lived condensations of plasma are
trapped along the star’s magnetic field lines, also near corotation
\citep{2022MNRAS.514.5465W}.  There is also a ``screen'' scenario, in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

In our view, the arguments seem best established for the dust clump
hypothesis \citep{2023MNRAS.518.4734S}.  This is primarily because the
prominence scenario lacks a mechanism for generating broadband
opacity, and the screen scenario seems inconsistent with the degree of
observed periodicity, and the lack of infrared excess.  However,
unambiguous evidence, for instance a spectroscopic detection of
silicates during a CQV ``dip'', has yet to be presented to confirm
which scenario is correct.

CQVs have been challenging to understand because they have been both
hard to discover and hard to characterize.   Discoverability is tied
to rarity: CQVs comprise about 1\% of the youngest 1\% of M-dwarfs
\citep{2018AJ....155..196R}.  Out of the millions of stars monitored
by K2 and TESS, about 50 CQVs have been reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2022AJ....163..144G,2023ApJ...945..114P}.
The known CQVs are correspondingly faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  This renders time-series spectroscopy
at high resolution technically impossible, despite its potential
utility in ruling between the models.

@@

A point on nomenclature.  CQVs have been called ``transient and
persistent flux dips'', ``scallop shells'', ``batwings'',
\citep{2017AJ....153..152S} ``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  The
CQVs should not be conflated with ``dippers'', which are classical T
Tauri stars with infrared excesses whose optical variability is linked
to obscuring inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  At the risk of
introducing yet another standard, we prefer a nomenclature that
reflects how, when observed over timescales of more than tens of
cycles, CQVs are never exactly periodic.  Their shapes evolve, usually
with some coherence timescale, similar to variability induced by
surface inhomogeneities such as starspots.  While the three-type
classification scheme proposed by \citet{2017AJ....153..152S} may
indeed provide some helpful visual distinctions amongst CQV
sub-classes, it seems likely that they are all explained by a single
underlying phenomenon, and so we opt to refer to them by a single
empirically descriptive name.

@@

		\vspace{-0.4cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{models.pdf}
		}
	\end{center}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
		{\it Top:} Phase-folded TESS light curves of three CQVs.  Each is
		stacked over one month.  Gray are raw 2-minute data; black bins to
		300 points per cycle.  Periods in hours are in the bottom right of
		each panel.  In order left-to-right, the objects are LP 12-502
		(TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and TIC
		425933644 (Sector~28).
		{\it Bottom:} Plausible cartoon models for the phenomenon.  The
		dust clump scenario seems most plausible, given
		the stability of the phenomenon, the lack of observed infrared
		excesses, and the challenge of producing broadband opacity variations
		with only ionized hydrogen in the prominence scenario.
	}
	\label{fig:f1}
\end{figure*}

@@

To assess the completeness of the resulting 120-second cadence data
that is the basis of this study, we cross-matched TIC8 \citep{2018AJ....156..102S}
against the Gaia DR2 point-source catalog \citep{2018A&A...616A...1G}.
We opted for Gaia DR2 rather than DR3 because the base catalog for
TIC8 was Gaia DR2, which facilitated a one-to-one crossmatch using the
Gaia source identifiers.  This exercise showed that for $T$$<$16
M-dwarfs, the TESS 2-minute data are roughly 50\% complete at
$\approx$50\,pc.  At $<$20\,pc, $\gtrsim$80\% of the $T$$<$16 M-dwarfs
have at least one sector of short-cadence data; at $>$100\,pc,
$\lesssim$10\% of such M-dwarfs have at least one sector of
short-cadence data.  Armed with this understanding, we then used our
cross-match between Gaia DR2 and TIC8 to select our stars of interest,
which we defined as stars with 120-second cadence TESS light curves
that satisfied
\begin{align}
  T &< 16 \quad&(\mathrm{Bright\ for\ TESS})\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only}) \\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})  \\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}),
\end{align}
for $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ the Gaia $G$-band
absolute magnitude, $\varpi_{\rm as}$ the parallax in units of
arcseconds, and a geometric distance $d$ defined by inverting the
parallax and ignoring any zero-point correction.  This selection
function includes dwarf stars later than spectral types of
$\approx$K6V \citep{2013ApJS..208....9P}.  The target sample therefore
includes \nstarssearched\ M-dwarfs and late-K dwarfs,
down to $T$$<$16 and out to $d$$<$150\,pc.

@@

Previous methods for finding CQVs have included visually
examining stars known to be in young clusters
\citep{2016AJ....152..114R,2017AJ....153..152S}, and automatically
flagging rapid rotators with a large number of strong Fourier
harmonics \citep{2019ApJ...876..127Z}.  The latter approach still requires
visual vetting, since ``stars with many Fourier harmonics''
is a designation that
includes objects such as eclipsing binaries or multiple stars
blended into a single photometric aperture.  In this work, we
implemented a new search approach based on counting the number of
sharp local minima in phase-folded light curves, while also using the
previously tested Fourier approach.  We applied these two search
techniques independently.

@@

{\bf TODO for Rahul or Saul: explain the approach, in a few
	paragraphs.  Was the SAP\_FLUX or PDCSAP used? etc. }

@@

\paragraph{Ages}
We estimated the stellar ages by making probabilistic spatial and
kinematic associations between the CQVs and known clusters in the
solar neighborhood.  For most stars in our sample, we did
this using BANYAN\,$\Sigma$
\citep{2018ApJ...856...23G}.\footnote{\url{https://github.com/jgagneastro/banyan_sigma},
git commit \texttt{394b486}} This algorithm calculates the probability
that a given star belongs to one of 27 young clusters (or
``assocations'') within 150\,pc of the Sun, by modeling the clusters
as multivariate Gaussians in 3-D position and 3-D velocity space.  We
used the Gaia DR2 sky positions, proper motions, and distances to
calculate the membership probabilities.  BANYAN\,$\Sigma$ in turn
analytically marginalizes over the radial velocity dimension.  The
probabilities returned by this procedure are qualitatively useful, but
we emphasize that they are quantitatively dubious due to the
non-Gaussian nature of most groups within the solar neighborhood
\citep[see e.g.][Figure~10]{2021ApJ...917...23K}.

@@

We determined the stellar effective temperature and radii through SED
fitting; we then estimated the masses by interpolating against
the PARSEC v1.2S models \citep{2012MNRAS.427..127B,2014MNRAS.444.2525C}.
%the ``magnetic'' models of \citet{2016A&A...593A..99F}.

For the SED fitting, we used \texttt{astroARIADNE}
\citep{2022MNRAS.513.2719V}.  Specifically, we adopted the ``BT-Settl''
stellar atmosphere models \citep{Allard2012} assuming the
\citet{2009ARA&A..47..481A} solar abundances, and the
\citet{2006MNRAS.368.1087B} water line lists.
The broadband magnitudes we considered included $GG_{\rm BP}G_{\rm
RP}$ from Gaia DR2, $Vgri$ from APASS, $JHK_{\rm S}$ from 2MASS, $riz$
from Pan-STARRS1, SDSS $riz$, and the WISE 1-2 passbands.  We
specifically omitted UV flux measurements to avoid biasing our fit
with any possible chromospheric UV excess.
\texttt{astroARIADNE} compares the measured broadband flux measurements
against pre-computed model grids, and by default fits for six
parameters:
$\{ T_{\rm eff}, R_\star, A_{\rm V}, \log g, [{\rm Fe/H}], d \}$.
The distance  prior is drawn from \citet{2021AJ....161..147B}.
The surface gravity and metallicity are generally unconstrained.
And finally, given our particular use-case, 
we assumed the following priors for the temperature, stellar size, and
extinction:
\begin{align}
  T_{\rm eff} / {\rm K}    &\sim \mathcal{N}(3000, 1000), \\
  R_\star / {\rm R}_\odot  &\sim \mathcal{T}_{\rm N}(0.5, 0.3, 0.1, 1.5), \\
  A_{\rm V} / {\rm mag}    &\sim \mathcal{U}(0, 0.2),
\end{align}
for $\mathcal{N}$ the Gaussian and $\mathcal{U}$ the uniform
distributions, and $\mathcal{T}(\mu, \sigma, a, b)$ a truncated normal
distribution with mean $\mu$, standard deviation $\sigma$, and lower
and upper bounds $a$ and $b$.
Using \texttt{Dynesty} \citep{2020MNRAS.493.3132S},
we statically sampled the posterior probability assuming the default
Gaussian likelihood, and set a stopping threshold of ${\rm d}\log
\mathcal{Z} < 0.01$, where $\mathcal{Z}$ denotes the evidence.

With the effective temperatures and stellar radii from the SED fit, we
then estimated the stellar masses by interpolating against the PARSEC
isochrones \citep[v1.2S][]{2014MNRAS.444.2525C}.  The need for models
that incorporate some form of correction for young, active M-dwarfs is
well-documented
\citep[e.g.][]{2012ApJ...756...47S,2015ApJ...804..146D,2016A&A...593A..99F,2020ApJ...891...29S}.
Plausible explanations for anomalous M-dwarf colors and sizes relative to
model predictions include star starspot coverage
\citep[e.g.][]{2017ApJ...836..200G}, and potentially incomplete line
lists \citep[e.g.][]{2013A&A...556A..15R}.  In the PARSEC models,
\citet{2014MNRAS.444.2525C} performed an empirical correction to the
temperature--opacity relation drawn from the BT-Settl model atmospheres,
in order to match observed masses and radii of younq eclipsing binaries.
This is sufficient for our purpose of estimating accurate stellar masses.
Given our observed $\{ \tilde{T}_{\rm eff}, \tilde{M}_\star, \tilde{t} \}$,
and approximating their uncertainties as Gaussian
$\sigma_{\tilde{T}_{\rm eff}}$, $\sigma_{\tilde{M}_\star}$ and
$\sigma_{\tilde{t}}$, we evaluate a distance $d$ between our observations
and any model PARSEC grid-point $\{ T_{\rm eff}, M_\star, t \}$ as
\begin{equation}
  d^2 = 
  \left( \frac{\tilde{T}_{\rm eff} - T_{\rm eff}}{\sigma_{\tilde{T}_{\rm eff}}} \right)^2
  +
  \left( \frac{\tilde{M}_{\star} - M_{\star}}{\sigma_{\tilde{M}_{\star}}} \right)^2
  +
  \left( \frac{\tilde{t} - t}{\sigma_{\tilde{t}}} \right)^2,
\end{equation}
in order to assign equal importance to each dimension.  The preferred
model mass is then one that minimizes this distance, and is quoted in
Table~\ref{tab:thetable}.

@@

\begin{figure*}[!t]
	\begin{center}
    \centering
    \includegraphics[width=0.96\textwidth]{402980664_S0018_120sec_cpvvetter.pdf}
		\vspace{-0.45cm}
		\caption{
      {\bf Validation plots used to label CQVs}.  The complete figure
      set, with one image per sector for each of \ncpvsfound\ objects
      is available online.
      Panels are as follows.
      {\it a)}: Phase-folded light curve; gray points are raw 2-minute
      data and black points are binned to 200 points per cycle.
      {\it b)}: Phase-dispersion minimization (PDM) periodogram.
      Dotted lines show up to the 10$^{\rm th}$ harmonic and
      subharmonic.
      {\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
      circles displayed for scale.
      {\it d)}: Cleaned light curve, binned to 20-minute cadence, in
      Barycentric TESS Julian Date (BTJD).
      {\it e)}: Phase-folded light curve, binned to 100 points per
      cycle.  The gray line denotes the automated spline-fit to the
      wrapped phase-folded light curve, and small gray triangles
      denote automatically identified local minima.
      {\it f)}: Phase-folded light curve at twice the peak period.
      {\it g)}: Phase-folded light curve at half the peak period.
      {\it h)}: Phase-folded time-series within the ``background''
      aperture defined in the SPOC light curves.
      {\it i)}: Phase-folded flux-weighted centroid in the column
      direction.
      {\it j)}: Phase-folded flux-weighted centroid in the row
      direction.
      {\it k)}: Gaia DR2 color--absolute magnitude diagram.     
      {\it l)}: Information from Gaia DR2, TIC8, and the automated
      dip-counting search pipeline.  ``Neighbors'', abbreviated
      ``nbhr'', are listed within apparent distances of 2 TESS pixels
      if $\Delta T$$<$2.5.
      {\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
      using positions, proper motions, and the parallax.
      }
		\label{fig:vet}
	\end{center}
\end{figure*}

\section{LP 12-502}
\label{app:lp}
